---
title: "StupidButtons"
author: "WangJY"
date: "2022-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

## clean data
```{r }
library(dplyr)
library(readxl)
library(tidyverse)
# nie<-read_xlsx("Data-NIE and WANG(1).xlsx",sheet = 1)
# pra<-read_xlsx("Mobile buttons.xlsx",sheet = 1)
# lbw<-read_xlsx("Data Lbw.xlsx",sheet = 1)
# alldata<-bind_rows(nie,pra,lbw)
# write.csv(alldata,row.names=F,"alldata.csv")
alldata<-read_csv("alldata.csv")
# alldata[alldata$Species=="Rhododendron leptothrium","Species"] <- "Rhododendron_leptothrium"
# write.csv(alldata,"alldata.csv",row.names=F)
summary(alldata)
```

## create species abundance matrix
```{r}
alldata <- alldata %>% mutate(ForestType=as.factor(ForestType),Species=as.factor(Species),Elevation=as.factor(Elevation),plotname=as.factor(paste0(ForestType,No.Plot,Elevation)))
summary(alldata)
length(unique(alldata$Species))
spabun<-as.data.frame(matrix(0,nrow=length(unique(alldata$plotname)),ncol=length(unique(alldata$Species))))
rownames(spabun)<-unique(alldata$plotname)
colnames(spabun)<-unique(alldata$Species)
summary(spabun)

for(i in 1:nrow(alldata)){
  spabun[as.character(alldata$plotname[i]),as.character(alldata$Species[i])]<-alldata$Number[i]
}
# write.csv(spabun,"SpeciesAbundance.csv")
```
We altogether have `r nrow(spabun)` sites, and `r ncol(spabun)` species. Thus the species abundance matrix consists `r nrow(spabun)` rows and `r ncol(spabun)` columns.

## look at trait
```{r}
library(tidyverse)
trait <- read_csv("dummy_trait.csv")
summary(trait)
```
Trait columns meaning: __LMA__ stands for Leaf mass per area (g m^2), __LL__ stands for Leaf lifespans (longevity, months), __Amass__ stands for Maximum photosynthetic rates per unit mass (nnmol g^-1 s^-1), __Rmass__ stands for Dark respiration rates per unit mass (nnmol g^-1 s^-1), __Nmass__ stands for Leaf nitrogen per unit mass (%), __Pmass__ stands for Leaf phosphorus per unit mass (%), __WD__ stands for Wood density (g cm^-3), __SM__ stands for Seed dry mass (mg).

```{r histogram for trait}
trait %>%
  pivot_longer(LMA:SM, names_to = "trait") %>%
  ggplot() + aes(x=value) +
  geom_histogram(position = "identity") +
  facet_wrap(~ trait, scale = "free")
# transforamtion
trait2 <- trait |>
  mutate(logLMA = log(LMA),
         logLL = log(LL),
         logAmass = log(Amass),
         logRmass = log(Rmass),
         logNmass = log(Nmass),
         logPmass = log(Pmass),
         logSM = log(SM)) |>
  dplyr::select(sp, logLMA, logLL, logAmass, logRmass, logNmass, logPmass, WD, logSM)

trait2 %>%
  pivot_longer(logLMA:logSM, names_to = "trait") %>%
  ggplot() + aes(x=value) +
  geom_histogram(position = "identity") +
  facet_wrap(~ trait, scale = "free")
# much normal now
```

## Diversity indices calculating
```{r shannon diversity}
diversity <- data.frame(plot=rownames(spabun))
diversity$shannon<-vegan::diversity(spabun, index = "shannon")

diversity$No.ind <- apply(spabun,1,sum)
```

```{r phylogenetic diversity}
library(picante)
library(FD)
phylo <- read.tree("dummy_tree.newick")
plot(phylo)
diversity[,c("PD","richness")] <- pd(spabun, phylo)
diversity[,"MNTD"] <-mntd(spabun, cophenetic(phylo))
diversity[,"MPD"] <- mpd(spabun, cophenetic(phylo))

```

```{r trait diversity}
trait_mat0 <- as.matrix(trait2[, -1])
rownames(trait_mat0) <- trait2$sp
# scale
trait_mat <- scale(trait_mat0)


tmp <- trait2 %>%
  filter(sp %in% colnames(spabun))

res_fd <- dbFD(trait_mat[colnames(spabun), ], spabun)

diversity$FRic <- res_fd$FRic
diversity$FDiv <- res_fd$FDiv
diversity$FDis <- res_fd$FDis
diversity$RaoQ <- res_fd$RaoQ
diversity$FEVe <- res_fd$FEve
```

## Diversity pattern across elevation and forest types

Now I have calculated all 
```{r see pattern}
library(stringr)
diversity$Elevation<-as.integer(str_sub(diversity$plot,start = str_length(diversity$plot)-2, end=-1))
diversity$forest <- as.factor(str_sub(diversity$plot,start = 1, end=str_length(diversity$plot)-4))
divcor <- cor(diversity[,2:12], method="pearson") 
corrplot::corrplot(divcor)

```
Didn't remove anything
```{r scatter plot}
diversity_long <- diversity %>% pivot_longer(shannon:RaoQ, names_to = "diversityindices")  %>% mutate(diversityindices=factor(diversityindices, levels=c("No.ind","richness","shannon","PD","MPD","MNTD","FRic","FEVe","FDiv","FDis","RaoQ")))
ggplot(diversity_long, aes(x = Elevation, y = value, colour = forest)) + geom_point() + geom_smooth(method = lm) +facet_wrap(~diversityindices,scales = "free",ncol=3) + theme_classic()
```
```{r linear regression}
library(lme4)
library(lmerTest)
library(performance)
ggplot(diversity_long,aes(x=value))+geom_histogram()+facet_wrap(~diversityindices,scale="free",ncol=3)

diversity %>% mutate(MNTD=log(MNTD),FRic=log(FRic),FEVe=(FEVe)^2,FDiv=(FDiv)^2,FDis=log(FDis),RaoQ=log(RaoQ)) %>%pivot_longer(shannon:RaoQ, names_to = "diversityindices")  %>% mutate(diversityindices=factor(diversityindices, levels=c("No.ind","richness","shannon","PD","MPD","MNTD","FRic","FEVe","FDiv","FDis","RaoQ"))) %>% ggplot()+ aes(x = value) + geom_histogram() +facet_wrap(~diversityindices,scales = "free",ncol=3) + theme_classic()
#good enough now

diversity.trans<-diversity %>% mutate(MNTD=log(MNTD),FRic=log(FRic),FEVe=(FEVe)^2,FDiv=(FDiv)^2,FDis=log(FDis),RaoQ=log(RaoQ))

diversity.trans %>% pivot_longer(shannon:RaoQ, names_to = "diversityindices")  %>% mutate(diversityindices=factor(diversityindices, levels=c("No.ind","richness","shannon","PD","MPD","MNTD","FRic","FEVe","FDiv","FDis","RaoQ"))) %>%ggplot()+aes(x = Elevation, y = value, colour = forest) + geom_point() + geom_smooth(method = lm) +facet_wrap(~diversityindices,scales = "free",ncol=3) + theme_classic()

diversity.trans$Elevation.sq <- (diversity.trans$Elevation)^2

##No. ind
indlm1 <- lm(No.ind~Elevation*forest, data=diversity.trans)
summary(indlm1) #signidicant
check_model(indlm1) # wrong!!!

indlm2 <- glm(No.ind~Elevation*forest, data=diversity.trans, family = "poisson")
summary(indlm2)
check_model(indlm2)
indlm3 <- glm(No.ind~Elevation, data=diversity.trans, family = "poisson")
summary(indlm3)
check_model(indlm3)

indlm4 <- update(indlm2,.~.+Elevation.sq*forest)
summary(indlm4) #not siignificant

AIC(indlm2,indlm3) # Although AIC lower, indlm3 diagnostic better

##richness
richnesslm1 <- lm(richness~Elevation*forest, data=diversity.trans)
summary(richnesslm1)
#not significant
check_model(richnesslm1)# wrong!!!
richnesslm2 <- glm(richness~Elevation*forest, data=diversity.trans, family = "poisson")
summary(richnesslm2)
#not significant
richnesslm3 <-update(richnesslm2,.~.+Elevation.sq*forest)
summary(richnesslm3)#not significant

##shannon
shannonlm1 <- lm(shannon~Elevation*forest,data=diversity.trans)
summary(shannonlm1)#not significant
shannonlm2 <-update(shannonlm1,.~.+Elevation.sq*forest)
summary(shannonlm2)#significant
check_model(shannonlm2) #fine

##PD
PDlm1 <- lm(PD~Elevation*forest, data=diversity.trans)
summary(PDlm1)
PDlm2 <- lm(PD~Elevation, data=diversity.trans)
summary(PDlm2)
PDlm3 <- update(PDlm1,.~.+Elevation.sq*forest)
summary(PDlm3) #not significant
AIC(PDlm1,PDlm2)
anova(PDlm1,PDlm2)
#just keep elevation

##MPD
MPDlm1 <- lm(MPD~Elevation*forest,data=diversity.trans)
summary(MPDlm1)#not significant
MPDlm2 <- update(MPDlm1,.~.+Elevation.sq*forest)
summary(MPDlm2)#not significant

##MNTD
MNTDlm1 <- lm(MNTD~Elevation*forest,data=diversity.trans)
summary(MNTDlm1)#not significant
MNTDlm2 <-update(MNTDlm1,.~.+Elevation.sq*forest)
summary(MNTDlm2)#not significant

##FRic
FRiclm1 <- lm(FRic~Elevation*forest,data=diversity.trans)
summary(FRiclm1) #not significant
FRiclm2 <- update(FRiclm1, .~.+Elevation.sq*forest)
summary(FRiclm2) # significant
check_model(FRiclm2) # diagnostic not good

##FEVe
FEVelm1 <- lm(FEVe~Elevation*forest,data=diversity.trans)
summary(FEVelm1)
#not significant
FEVelm2<-update(FEVelm1, .~.+Elevation.sq*forest)
summary(FEVelm2)# not significant

##FDiv
FDivlm1 <- lm(FDiv~Elevation*forest,data=diversity.trans)
summary(FDivlm1)# significant
check_model(FDivlm1) # not good enough but make sense
FDivlm2 <- update(FDivlm1,.~.+Elevation.sq*forest)
summary(FDivlm2) # not significant



##FDis

FDislm1 <- lm(FDis~Elevation*forest,data=diversity.trans)
summary(FDislm1)# significanyt
FDislm2 <-update(FDislm1, .~.+Elevation.sq*forest)
summary(FDislm2) # very very significant
check_model(FDislm2) #good


##RaoQ

RaoQlm1 <-lm(RaoQ~Elevation*forest,data=diversity.trans)
summary(RaoQlm1) # not significance
RaoQlm2 <- update(RaoQlm1,.~.+Elevation.sq*forest)
summary(RaoQlm2) # significant
check_model(RaoQlm2)# not good

```

Significant models: indlm3, shannonlm2, PDlm2, FDivlm1, FDislm2, last two transformed

```{r predictioin using lm}
indlm3$call
check_overdispersion(indlm3) #no overdispersion

indnew <- data.frame(Elevation=rep(seq(500,900,10),2),forest=rep(c("Primary","Secondary"),each=41))
indnew$Elevation.sq <-(indnew$Elevation)^2

preds <- predict(indlm3, newdata=indnew,type= 'link', se.fit=T)
indnew$No.ind <- exp(preds$fit)
indnew$indup <- exp(preds$fit+preds$se.fit*1.96)
indnew$indlw <- exp(preds$fit-preds$se.fit*1.96)

indp <- ggplot(diversity,aes(x=Elevation,y=No.ind,color=forest))+geom_point()+geom_smooth(data=indnew,aes(x=Elevation, y=No.ind, ymin=indlw, ymax=indup),stat='identity')+theme_classic()



shannonlm2$call

shannonnew <- data.frame(Elevation=rep(seq(500,900,10),2),forest=rep(c("Primary","Secondary"),each=41))
shannonnew$Elevation.sq <-(shannonnew$Elevation)^2

preds <- predict(shannonlm2, newdata=shannonnew, se.fit=T)
shannonnew$shannon <- preds$fit
shannonnew$shannonup <- preds$fit+preds$se.fit*1.96
shannonnew$shannonlw <- preds$fit-preds$se.fit*1.96

shannonp <- ggplot(diversity,aes(x=Elevation,y=shannon,color=forest))+geom_point()+geom_smooth(data=shannonnew,aes(x=Elevation, y=shannon, ymin=shannonlw, ymax=shannonup),stat='identity')+theme_classic()



PDlm2$call
summary(PDlm2)
PDnew <- data.frame(Elevation=rep(seq(500,900,10),2),forest=rep(c("Primary","Secondary"),each=41))
PDnew$Elevation.sq <-(PDnew$Elevation)^2

preds <- predict(PDlm2, newdata=PDnew, se.fit=T)
PDnew$PD <- preds$fit
PDnew$PDup <- preds$fit+preds$se.fit*1.96
PDnew$PDlw <- preds$fit-preds$se.fit*1.96

PDp <- ggplot(diversity,aes(x=Elevation,y=PD,color=forest))+geom_point()+geom_smooth(data=PDnew,aes(x=Elevation, y=PD, ymin=PDlw, ymax=PDup),stat='identity')+theme_classic()


FDivlm1$call
summary(FDivlm1)
FDivnew <- data.frame(Elevation=rep(seq(500,900,10),2),forest=rep(c("Primary","Secondary"),each=41))
FDivnew$Elevation.sq <-(FDivnew$Elevation)^2

preds <- predict(FDivlm1, newdata=FDivnew, se.fit=T)
FDivnew$FDiv <- sqrt(preds$fit)
FDivnew$FDivup <- sqrt(preds$fit+preds$se.fit*1.96)
FDivnew$FDivlw <- sqrt(preds$fit-preds$se.fit*1.96)

FDivp <- ggplot(diversity,aes(x=Elevation,y=FDiv,color=forest))+geom_point()+geom_smooth(data=FDivnew,aes(x=Elevation, y=FDiv, ymin=FDivlw, ymax=FDivup),stat='identity')+theme_classic()


FDislm2$call
summary(FDislm2)
FDisnew <- data.frame(Elevation=rep(seq(500,900,10),2),forest=rep(c("Primary","Secondary"),each=41))
FDisnew$Elevation.sq <-(FDisnew$Elevation)^2

preds <- predict(FDislm2, newdata=FDisnew, se.fit=T)
FDisnew$FDis <- exp(preds$fit)
FDisnew$FDisup <- exp(preds$fit+preds$se.fit*1.96)
FDisnew$FDislw <- exp(preds$fit-preds$se.fit*1.96)

FDisp <- ggplot(diversity,aes(x=Elevation,y=FDis,color=forest))+geom_point()+geom_smooth(data=FDisnew,aes(x=Elevation, y=FDis, ymin=FDislw, ymax=FDisup),stat='identity')+theme_classic()




library(patchwork)
indp+shannonp+PDp+FDivp+FDisp+plot_layout(ncol = 2,guides="collect")
```




