---
title: "StupidButtons"
author: "WangJY"
date: "2022-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

## clean data
```{r }
library(dplyr)
library(readxl)
library(tidyverse)
# nie<-read_xlsx("Data-NIE and WANG(1).xlsx",sheet = 1)
# pra<-read_xlsx("Mobile buttons.xlsx",sheet = 1)
# lbw<-read_xlsx("Data Lbw.xlsx",sheet = 1)
# alldata<-bind_rows(nie,pra,lbw)
# write.csv(alldata,row.names=F,"alldata.csv")
alldata<-read_csv("alldata.csv")
# alldata[alldata$Species=="Rhododendron leptothrium","Species"] <- "Rhododendron_leptothrium"
# write.csv(alldata,"alldata.csv",row.names=F)
summary(alldata)
```

## create species abundance matrix
```{r}
alldata <- alldata %>% mutate(ForestType=as.factor(ForestType),Species=as.factor(Species),Elevation=as.factor(Elevation),plotname=as.factor(paste0(ForestType,No.Plot,Elevation)))
summary(alldata)
length(unique(alldata$Species))
spabun<-as.data.frame(matrix(0,nrow=length(unique(alldata$plotname)),ncol=length(unique(alldata$Species))))
rownames(spabun)<-unique(alldata$plotname)
colnames(spabun)<-unique(alldata$Species)
summary(spabun)

for(i in 1:nrow(alldata)){
  spabun[as.character(alldata$plotname[i]),as.character(alldata$Species[i])]<-alldata$Number[i]
}
# write.csv(spabun,"SpeciesAbundance.csv")
```
We altogether have `r nrow(spabun)` sites, and `r ncol(spabun)` species. Thus the species abundance matrix consists `r nrow(spabun)` rows and `r ncol(spabun)` columns.

## look at trait
```{r}
library(tidyverse)
trait <- read_csv("dummy_trait.csv")
summary(trait)
```
Trait columns meaning: __LMA__ stands for Leaf mass per area (g m^2), __LL__ stands for Leaf lifespans (longevity, months), __Amass__ stands for Maximum photosynthetic rates per unit mass (nnmol g^-1 s^-1), __Rmass__ stands for Dark respiration rates per unit mass (nnmol g^-1 s^-1), __Nmass__ stands for Leaf nitrogen per unit mass (%), __Pmass__ stands for Leaf phosphorus per unit mass (%), __WD__ stands for Wood density (g cm^-3), __SM__ stands for Seed dry mass (mg).

```{r histogram for trait}
trait %>%
  pivot_longer(LMA:SM, names_to = "trait") %>%
  ggplot() + aes(x=value) +
  geom_histogram(position = "identity") +
  facet_wrap(~ trait, scale = "free")
# transforamtion
trait2 <- trait |>
  mutate(logLMA = log(LMA),
         logLL = log(LL),
         logAmass = log(Amass),
         logRmass = log(Rmass),
         logNmass = log(Nmass),
         logPmass = log(Pmass),
         logSM = log(SM)) |>
  dplyr::select(sp, logLMA, logLL, logAmass, logRmass, logNmass, logPmass, WD, logSM)

trait2 %>%
  pivot_longer(logLMA:logSM, names_to = "trait") %>%
  ggplot() + aes(x=value) +
  geom_histogram(position = "identity") +
  facet_wrap(~ trait, scale = "free")
# much normal now
```

## Diversity indices calculating
```{r shannon diversity}
diversity <- data.frame(plot=rownames(spabun))
diversity$shannon<-vegan::diversity(spabun, index = "shannon")

diversity$No.ind <- apply(spabun,1,sum)
```

```{r phylogenetic diversity}
library(picante)
library(FD)
phylo <- read.tree("dummy_tree.newick")
plot(phylo)
diversity[,c("PD","richness")] <- pd(spabun, phylo)
diversity[,"MNTD"] <-mntd(spabun, cophenetic(phylo))
diversity[,"MPD"] <- mpd(spabun, cophenetic(phylo))
```

```{r trait diversity}
trait_mat0 <- as.matrix(trait2[, -1])
rownames(trait_mat0) <- trait2$sp
# scale
trait_mat <- scale(trait_mat0)


tmp <- trait2 %>%
  filter(sp %in% colnames(spabun))

res_fd <- dbFD(trait_mat[colnames(spabun), ], spabun)

diversity$FRic <- res_fd$FRic
diversity$FEVe <- res_fd$FEve
diversity$FDiv <- res_fd$FDiv
diversity$FDis <- res_fd$FDis
diversity$RaoQ <- res_fd$RaoQ
```

## Diversity pattern across elevation and forest types

Now I have calculated all 
```{r see pattern}
library(stringr)
diversity$Elevation<-as.integer(str_sub(diversity$plot,start = str_length(diversity$plot)-2, end=-1))
diversity$forest <- as.factor(str_sub(diversity$plot,start = 1, end=str_length(diversity$plot)-4))
divcor <- cor(diversity[,2:12], use="pairwise.complete.obs", method="pearson") 
corrplot::corrplot(divcor)

```
Remove High correlation pairs: RaoQ, FDis
```{r scatter plot}
diversity_long <- diversity %>% select(-RaoQ,-FDis) %>% pivot_longer(shannon:FDiv, names_to = "diversityindices") 
ggplot(diversity_long, aes(x = Elevation, y = value, colour = forest)) + geom_point() + geom_smooth(method = lm) +facet_wrap(~diversityindices,scales = "free") + theme_classic()
```
```{r linear regression}

```
